---
title: "Part 1 - Data Cleaning by Correction"
output: html_document
---
###*UNI:sv2414*
##**NOTE**:
##**1.This program writes the final output to file. Check "Part1_output.csv" for all values described here**
##**2.The actual code can be viewed in the Rmd file "Part 1 - Data Cleaning by Correction.Rmd"**

###Check if packages are installed, install if required, and load
```{r, echo=FALSE}
if (!require(plyr)) install.packages("plyr")
library(plyr)
if (!require(ggplot2)) install.packages("ggplot2")
library(ggplot2)
```

###Import CSV file containing wind power generation data
###*Summary of "Windpower" data frame is shown below:*
```{r, echo=FALSE}
windpowercsv<-"WIND_VXE_2013_ORIGINAL.csv"
windpower<-read.csv(windpowercsv)
summary(windpower)
```

###Find initial number of rows
```{r, echo=FALSE}
orig_size=nrow(windpower)
sprintf("There are %d rows.",orig_size)
```

###Remove columns I don't need
###*Updated summary:*
```{r, echo=FALSE}
windpower<-windpower[-c(2:15,17,18)]
summary(windpower)
```

###Rename columns
###*Updated summary:*
```{r, echo=FALSE}
names(windpower)[1]<-"DateTime"
names(windpower)[2]<-"AvgWindSpeed"
names(windpower)[3]<-"MeterReading"
summary(windpower)
```

###Convert DateTime to Date-Time values (useful for Part 2)
```{r, echo=FALSE}
windpower$DateTime<-as.POSIXct(windpower$DateTime,format="%m/%d/%Y %H:%M")
```

###Check for negative values
```{r, echo=FALSE}
negstat=0
if(summary(windpower$AvgWindSpeed)[1]>=0) {
  print("No negative values found in Wind Speed values.")
  } else {
  print("Negative values found in Wind Speed values.")
  negstat=1
  }
if(summary(windpower$MeterReading)[1]>=0) {
  print("No negative values found in Meter Reading values.")
  } else {
  print("Negative values found in Meter Reading values.")
  negstat=1
  }
```

###Convert negative values, if any, to NA
```{r, echo=FALSE}
if(negstat==1){
  windpower$AvgWindSpeed[windpower$AvgWindSpeed<0]<-NA
  windpower$MeterReading[windpower$MeterReading<0]<-NA
}
```

###Convert missing and negative Wind Speed values to 0
###(This is because wind speeds are assumed to be faulty and can be corrected (need not be removed))
```{r, echo=FALSE}
windpower$AvgWindSpeed[which(is.na(windpower$AvgWindSpeed))]<-0
```

###Print total number of faulty values found in preliminary cleaning
```{r, echo=FALSE}
sprintf("Total number of missing or negative values is %d", sum(is.na(windpower)))
print("These are contained in the following rows:")
windpower[which(is.na(windpower$MeterReading)),]
print("Omitting these...")
windpower<-na.omit(windpower)
sprintf("After omission of missing values, there are now %d observations remaining.", nrow(windpower))
```

###Add new column for kWh delivered in 10 minutes
###*Summary of "tenminkwh" column is shown below:*
```{r, echo=FALSE}
windpower$tenminkwh<-0
for(i in 2:nrow(windpower)){
  windpower$tenminkwh[i]=windpower$MeterReading[i]-windpower$MeterReading[i-1]
  #Since the meter resets after reaching maximum reading
  if(windpower$MeterReading[i]<windpower$MeterReading[i-1]){
    windpower$tenminkwh[i]=9999999-windpower$MeterReading[i-1]+windpower$MeterReading[i]
  }
}
summary(windpower$tenminkwh)
```

###Remove power generated values greater than rated capacity
```{r, echo=FALSE}
windpower$tenminkwh[windpower$tenminkwh>850*7/6]<-NA
print("Omitting these...")
windpower<-na.omit(windpower)
sprintf("After omission of missing values, there are now %d observations remaining.", nrow(windpower))
```

###Import CSV file containing Manufacturer's PowerCurve
###*Summary of "mpc" data frame is shown below:*
```{r, echo=FALSE}
manpc<-"v52-850KW-power-curve.csv"
mpc<-read.csv(manpc)
summary(mpc)
```

###Rename columns
###*Updated summary:*
```{r, echo=FALSE}
names(mpc)[1]<-"WindSpeed"
names(mpc)[2]<-"Power"
summary(mpc)
```

###WE ARE ASSUMING METER READINGS ARE CORRECT AND WIND VALUES *MAY* BE FAULTY
###Add new column for 10min generation according to MPC, Betz Limit, Kinetic Energy of Wind
###This is just to examine inconsistencies in a plot
###*Summary of "tenminmpcurve", "tenminbetz", "tenminKEwind" columns are shown below:*
```{r, echo=FALSE}
interp=splinefun(mpc$WindSpeed, mpc$Power,method = "natural")
windpower$tenminmpcurve<-interp(windpower$AvgWindSpeed)*(7/6)
windpower$tenminmpcurve[windpower$tenminmpcurve<0]<-0
windpower$tenminbetz<-0.593*0.5*1.225*2124*((windpower$AvgWindSpeed)^3)*(7/6000)
windpower$tenminKEwind<-1*0.5*1.225*2124*((windpower$AvgWindSpeed)^3)*(7/6000)
summary(windpower$tenminmpcurve)
summary(windpower$tenminbetz)
summary(windpower$tenminKEwind)

```

###Plot to see inconsistencies
```{r, echo=FALSE}
plot1<-ggplot(windpower, aes(x=AvgWindSpeed))
plot1+geom_point(aes(y=windpower$tenminkwh),colour="yellow",size=1)+geom_line(aes(y=windpower$tenminmpcurve),colour="blue")+geom_line(aes(y=windpower$tenminbetz),colour="red")+geom_line(aes(y=windpower$tenminKEwind),colour="black")+ylim(0,1000)+ylab("Ten Min Generation (kWh)")+xlab("Average Wind Speed (m/s)")+ggtitle("Ten Min Generation Data After PRELIMINARY Cleaning\nYellow=Data | Blue=MPC | Red=BetzLimit | Black=KEinWind")
```
---
###Everything above the MPC should be moved to the MPC
###Add a column to calculate equivalent power in KW for ten minute intervals
###*Summary of "eqPower" column is shown below:*
```{r, echo=FALSE}
windpower$eqPower=windpower$tenminkwh*(6/7)
summary(windpower$eqPower)
```

###Import CSV file containing Manufacturer's PowerCurve sorted by Power and cleaned
###(This is because we are interpolating for wind speed based on power)
###*Summary of "mpc2" data frame is shown below:*
```{r, echo=FALSE}
manpc2<-"mpc_powersort.csv"
mpc2<-read.csv(manpc2)
summary(mpc2)
```

###Rename columns
###*Updated summary:*
```{r, echo=FALSE}
names(mpc2)[1]<-"Power"
names(mpc2)[2]<-"WindSpeed"
summary(mpc2)
```

###Add a column to interpolate for wind values based on MPC
###*Summary of "mpcWind" column is shown below:*
```{r, echo=FALSE}
interp2=splinefun(x = mpc2$Power,y = mpc2$WindSpeed,method = "natural")
windpower$mpcWind=interp2(windpower$eqPower)
windpower$mpcWind[windpower$mpcWind<0]<-0
summary(windpower$mpcWind)
```

###Add a new column for final cleaned wind speed values
###Compare measured and interpolated wind values, and correct
###*Summary of "finalWSvalue" column is shown below:*
```{r, echo=FALSE}
windpower$finalWSvalue<-0
for (i in 1:nrow(windpower)){
  if(windpower$AvgWindSpeed[i]<windpower$mpcWind[i]){
    windpower$finalWSvalue[i]=windpower$mpcWind[i]
  }
  else{
    windpower$finalWSvalue[i]=windpower$AvgWindSpeed[i]
  }
}
summary(windpower$finalWSvalue)
```

###Plot to see cleaned results
```{r, echo=FALSE}
plot2<-ggplot(windpower, aes(x=finalWSvalue))
plot2+geom_point(aes(y=windpower$tenminkwh),colour="yellow",size=1)+geom_line(aes(x=windpower$AvgWindSpeed,y=windpower$tenminmpcurve),colour="blue")+geom_line(aes(x=windpower$AvgWindSpeed,y=windpower$tenminbetz),colour="red")+geom_line(aes(x=windpower$AvgWindSpeed,y=windpower$tenminKEwind),colour="black")+ylim(0,1000)+ylab("Ten Min Generation (kWh)")+xlab("Average Wind Speed (m/s)")+ggtitle("Ten Min Generation Data After FINAL Cleaning\nYellow=Data | Blue=MPC | Red=BetzLimit | Black=KEinWind")
```
---
###This is satisfactory; everything is on or below the MPC!
###Add new rows for actual and uncurtailed generation
###*Summary of "ActualGenerationkWh", "UncurtailedGenerationkWh" columns are shown below:*
```{r, echo=FALSE}
windpower$ActualGenerationkWh=windpower$tenminkwh
windpower$UncurtailedGenerationkWh=interp(windpower$finalWSvalue)*(7/6)
summary(windpower$ActualGenerationkWh)
summary(windpower$UncurtailedGenerationkWh)
```

###Calculate total annual actual and uncurtailed generation
```{r, echo=FALSE}
tot_act_gen=sum(windpower$ActualGenerationkWh)
tot_unc_gen=sum(windpower$UncurtailedGenerationkWh)
sprintf("ANNUAL GENERATION IS %.2fkWh",tot_act_gen)
sprintf("ANNUAL UNCURTAILED GENERATION IS %.2fkWh",tot_unc_gen)
```

###Calculate total possible generation at nameplate capacity (850kW)
```{r, echo=FALSE}
tot_np=850*7*8760
```

###Calculate actual and uncurtailed Capacity Factors
```{r, echo=FALSE}
act_cp=tot_act_gen*100/tot_np
unc_cp=tot_unc_gen*100/tot_np
sprintf("ACTUAL CAPACITY FACTOR IS %.1fpc",act_cp)
sprintf("UNCURTAILED CAPACITY FACTOR IS %.1fpc",unc_cp)
```

###Add a column for Kinetic Energy in wind at cleaned values of wind speeds
###*Summary of "KEinWind" column is shown below:*
```{r, echo=FALSE}
windpower$KEinWind=1*0.5*1.225*2124*((windpower$finalWSvalue)^3)*(7/6000)
summary(windpower$KEinWind)
```

###Add a column for Turbine Efficiency
###*Summary of "TurbineEfficiency" column is shown below:*
```{r, echo=FALSE}
windpower$TurbineEfficiency=windpower$ActualGenerationkWh*100/windpower$KEinWind
summary(windpower$TurbineEfficiency)
```

###Calculate average Turbine Efficiency for the year
```{r, echo=FALSE}
avg_te=sum(windpower$ActualGenerationkWh)*100/sum(windpower$KEinWind)
sprintf("Average Turbine Efficiency for the year is %.2fpc",avg_te)
```

###Plot turbine efficiency and compare to Betz Limit
```{r, echo=FALSE}
plot3<-ggplot(windpower, aes(x=finalWSvalue))
plot3+geom_point(aes(y=windpower$TurbineEfficiency),colour="yellow",size=1)+geom_line(aes(y=avg_te),colour="blue")+geom_line(aes(y=59.3),colour="red")+ylim(0,100)+ylab("Turbine Efficiency (%)")+xlab("Average Wind Speed (m/s)")+ggtitle("Turbine Efficiency\nYellow=EfficiencyData | Blue=AverageEfficiency | Red=BetzLimit")
```
---
###It looks to be right!
###Find final number of rows
```{r, echo=FALSE}
fin_size=nrow(windpower)
sprintf("Final number of rows (after all cleaning) is %d",fin_size)
sprintf("Total number of rows omitted is %d",orig_size-fin_size)
```

###Write data to file (to be used in Part 2)
```{r, echo=FALSE}
write.csv(windpower,"Part1_output.csv")
sprintf("Printed to: Part1_output.csv")
```
##**FINAL COMMENTS:**
##**1.The actual, uncurtailed capacity factors and the turbine efficiency is found to be as expected (40.7%, 52.4%. 28.3%).**
##**2. The number of observations after cleaning is nearly the same as before. Negligible difference.**